---
title: "Assignment 1+2"
author: "Grace Levin, Sam Lewis, Naomi Robalino"
date: "10/28/2021"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

## Asignment 1

## 1. Research Question

How well does the increase in rent prices predict the number of people experiencing homelessness in the United States on the state level? 

## 2. Prior Research

At first, we wanted to know the relationship between types of affordable housing and homelessness rates. However, according to Petry et. al. (2021) people at high risk of experiencing homelessness were likely to re-enter homelessness regardless of their housing type; they found those at high risk of experiencing homeslessness depended on their structural vulnerabilities. We instead decided to focus on larger scale forces such as the housing market and median rent prices in order to see how it affects the homelessness rate in the U.S. at the state level. Additionally, according to the U.S. Interagency Council on Homelessness (2021), households with higher levels of housing stability and access to housing affordability have a lower likelihood of entering homelessness, so we decided to look at the percent increase in rents as well. 

## 3. Data

The sample set for this study is all states in the United States. The analysis will include the following variables:

Categorical:

* Region the U.S. state in is (U.S (American Community Survey 2019) [region]
* Majority party in the state legislature ("State Legislative Elections, 2021 - Ballotpedia" 2021) [party_stateleg]

Continuous:

* Number of people experience poverty (American Community Survey 2019) 
[povertyE]
* Median gross rent (American Community Survey 2019) [median_rent]
* Percent Increase in rents since 2001 (American Community Survey 2019) [perincrease_rent]
* Total number of people experiencing homelessness (HUD 2018 Annual Homelessness Assessment Report) [homelessness]
* Total number of beds available in homeless shelters (PIT and HIC Data Since 
2007, 2021) [bed_counts_per_state]

This dataset includes 50 states and the District of Columbia. 



## 4. References

"National And State Housing Fact Sheets & Data". 2019. Center On Budget And Policy Priorities. https://www.cbpp.org/research/housing/national-and-state-housing-fact-sheets-data.

Petry, Laura, Chyna Hill, Phebe Vayanos, Eric Rice, Hsun-Ta Hsu, and Matthew Morton. “Associations Between the Vulnerability Index-Service Prioritization Decision Assistance Tool and Returns to Homelessness Among Single Adults in the United States.” Cityscape 23, no. 2 (2021): 293–324. https://www.jstor.org/stable/27039964.

"PIT And HIC Data Since 2007". 2021. Hudexchange.Info. https://www.hudexchange.info/resource/3031/pit-and-hic-data-since-2007/.

"State Legislative Elections, 2021 - Ballotpedia". 2021. Ballotpedia. https://ballotpedia.org/State_legislative_elections,_2021.

2021. Usich.Gov. https://www.usich.gov/resources/uploads/asset_library/Housing-Affordability-and-Stablility-Brief.pdf.

United States Census Bureau. American Community Survey, 5-year estimates. 2019.



## 5. Load Packages
First we loaded the following packages:

```{r}

library(osmextract)
library(tidyverse)
library(sf)
library(ggthemes)
library(ggspatial)
library(tigris)
library(wesanderson)
library(tidytransit)
library(tidycensus)
library(dplyr)
library(stars)
library(readxl)
library(knitr)
library(gridExtra)

```

## 6. Load Data

Then, we imported the American Community Census data recorded in 2019 (from the year 2015-2019) and the census data

```{r}
census_api_key("32af35a06c38b3b5eaf52448f34555621ee04370", install=TRUE, overwrite = TRUE)
readRenviron("~/.Renviron")

acs_vars <- load_variables(2019, "acs1")
```

Then we loaded the state poverty levels from the 2019 ACS data.

```{r}
state_poverty <- get_acs(geography = "state",
                         year = 2019,
                          output = "wide",
                          variables = c(poverty = 'B05010_002', pop = 'B01003_001'))
  
```


Then we loaded the state population from the 2019 ACS data.

```{r}

state_pop <- get_acs(geography = "state",
                         year = 2019,
                          output = "wide",
                          variables = c(population = 'B01003_001'))

```



Then, we loaded state housing facts.

```{r}
state_housing <- read_csv(file = 'statehousingfacts.csv')
```

Then we loaded bed counts per state.

```{r}
state_shelter <- read_csv(file = 'state_bed.csv') %>%
  rename (bed_counts_per_state = beds)
```

## 7. Join Data

Lastly, we joined the state poverty, state housing, state shelter.

```{r}
data <- left_join(state_poverty, state_housing, by = 'NAME') %>% 
  left_join(state_shelter)

data <- select(data, -popM, -povertyM)

knitr::kable(data)

```

---
title: "Assignment 2"
---

## Calculate Continuous variables (Quartiles, SD,

```{r}
poverty_t_test <- t.test(data$povertyE)
popE_t_test <- t.test(data$popE)
median_rent_t_test <- t.test(data$median_rent)
perincrease_rent_t_test <- t.test(data$perincrease_rent)
homelessness_t_test <- t.test(data$homelessness)
bed_counts_per_state_t_test <- t.test(data$bed_counts_per_state)

poverty_quartiles <- quantile(data$povertyE, na.rm = TRUE)
popE_quartiles <- quantile(data$popE, na.rm = TRUE)
median_rent_quartiles <- quantile(data$median_rent, na.rm = TRUE)
perincrease_rent_quartiles <- quantile(data$perincrease_rent, na.rm = TRUE)
homelessness_quartiles <- quantile(data$homelessness, na.rm = TRUE)
bed_counts_per_state_quartiles <- quantile(data$bed_counts_per_state, na.rm = TRUE)


poverty_st_dev <- sd(data$povertyE, na.rm = TRUE)
popE_st_dev <- sd(data$popE, na.rm = TRUE)
median_rent_st_dev <- sd(data$median_rent, na.rm = TRUE)
perincrease_rent_st_dev <- sd(data$perincrease_rent, na.rm = TRUE)
homelessness_st_dev <- sd(data$homelessness, na.rm = TRUE)
bed_counts_per_state_st_dev <- sd(data$bed_counts_per_state, na.rm = TRUE)


```


```{r}

poverty_hist <- ggplot(data) +
  geom_histogram(aes(x = povertyE),
                 bins = 30)

popE_hist <- ggplot(data) +
  geom_histogram(aes(x = popE),
                 bins = 30)

median_rent_hist <- ggplot(data) +
  geom_histogram(aes(x = median_rent),
                 bins = 30)

perincrease_rent_hist <- ggplot(data) +
  geom_histogram(aes(x = perincrease_rent),
                 bins = 30)

homelessness_hist <- ggplot(data) +
  geom_histogram(aes(x = homelessness),
                 bins = 30)

bed_counts_per_state_hist <- ggplot(data) +
  geom_histogram(aes(x = bed_counts_per_state),
                 bins = 30)

```


## Poverty
```{r}

poverty_t_test 
poverty_st_dev
poverty_hist

```
## Population
```{r}
popE_t_test 
popE_st_dev
popE_hist
```
## Median Rent
```{r}
median_rent_t_test 
median_rent_st_dev
median_rent_hist
```

## Percent Increase in Rent
```{r}
perincrease_rent_t_test 
perincrease_rent_st_dev
perincrease_rent_hist
```

## I
```{r}
homelessness_t_test 
homelessness_st_dev
homelessness_hist
```


## Homeless beds
```{r}
bed_counts_per_state_t_test 
bed_counts_per_state_st_dev
bed_counts_per_state_hist

```

